### 2021-01-07 paper/main.tex
restructure, point form intro notes

\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\newcommand\myworries[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Inverting Colors to save my eye balls
%\usepackage{xcolor}
%\pagecolor[rgb]{0.1,0.12,0.15}
%\color[rgb]{.9,.9,.95}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{acmauthoryear}

\begin{document}

\title{Shape Matching Element Method}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{castle.png}
  \caption{tmp}
  \Description{}
  \label{fig:teaser}
\end{teaserfigure}

\maketitle

\input{sections/intro}
\input{sections/related}
\input{sections/results}
\input{sections/conclusion}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

Dump of some equations
\begin{align*}
    \mathbf{M(X)} = (X_0 - \bar{X}_0, X_1 - \bar{X}_1)\\
    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i \\
    \bar{\mathbf{X}} = \frac{1}{n}\sum_i^n X_i \\
\end{align*}

\end{document}