### 2021-01-07 paper/sections/intro.tex

\section{Introduction}

% Modelling -> Animation -> Rendering
% Use physics for animation which requires conversion from modelling to simulation representation
% for volumetric objects this conversion invovles identififyin and meshing the interior of the modelled object and that's hard
% especially difficult if object surface is described by higher-order boundary representations. 
% Ideally we would have a more direct method of translating our surface model into a volumetric, simulatable form
%
% NURBs  are an oft-used higher order modelling primitive popular for CAD and Computer Graphics for which nice modelling tools exist
% Rhinocerous 3D and Fusion 360.
% Despite lots of work performing physics-based animation on NURBS models is hard
% robustly uilding a volumetric polygonal mesh that simulataneously maintains a corresonpondence to the model surface is an open-research problem
% Isogeometric approaches which directly use the NURBs primitives as simulation variables require additional background grids, complex cutcell algorithms
% and often must rely on soft boundary conditions.
%
% Want to say there is an advantage to avoiding constructing volumetric representations because it costs time and adds complexity
% that;s our goal to provide an alternative, the first truly meshless approach to IGA which maintains a tightt mapping to the boundary
% In this paper we present a new finite element method for isogeometric, physics-based animation of NURBs models.
% Crucially our method is boundary only, and requires no volumetric meshes (thus avoiding their inherent complications), grids or cutcells of any kind
% Our method does not need an explicit labelling of the inside/outide of the simulated model
% Does not require nurbs patches in an object to be explicitly joined
% Input is just the raw NURBs model with no additional information or annotations
% The resulting simulation algorithm is compatiable with nonlinear continuum mechanics material models as well as 
% standard methods for contact resolution. 
%
% crucially because we broadly target animation not necessarily simulation for engineering or manufacturing we don't require that models
% satisfy rigourous geometric requirements common for these applications
%
% Our approach relies on the recently developed Virtual Element Method for solving PDEs on domains tiled with arbitrary polygons.
% By establsihing the connection between VEM and the well-known shape matching simulation algorithm we derive equations of motion sfor 
% an arbitrary NURBS model using Lagrangian Mechanics. We avoid using volumetric data structures for integration by demonstrating that 
% raycasting provides a fast and reliable alternative. In combination, our contributions amount to the first algorithm for 
% isogeometric elastodynamics of volumetric structures that is truly meshless (introducing no auxilillary volumetric data structures) 
% which serves to close the considerable gap betweeen modelling and physics-based animation. 
%
% Desiderata for a sucessful algorithm for PBA 
% No volumetric mesh creation
% Support for Nonlinear Constitutive Models
% Support for standard Collision 
% DOFS on the Boundary (useful for boundary conditions, contact)
% Automatic determination of model volume
% Non-manufacturable models 
% Code Available

%limitations
% sometimes you want some volumetric dofs to better resolve internal behaviour