### 2021-01-16 paper/main.tex
starting methods. draft of method overview section

\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\input{sections/pre.tex}

\newcommand\myworries[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Inverting Colors to save my eye balls
%\usepackage{xcolor}
%\pagecolor[rgb]{0.1,0.12,0.15}
%\color[rgb]{.9,.9,.95}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{acmauthoryear}

\begin{document}

\title{The Shape Matching Element Method for Meshless Animation of NURBs Models}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{castle.png}
  \caption{tmp}
  \Description{}
  \label{fig:teaser}
\end{teaserfigure}

\maketitle

\input{sections/intro}
\input{sections/related}
\input{sections/methods}

\input{sections/results}
\input{sections/conclusion}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}