### 2021-01-26 paper/main.tex
crummy abstract

\documentclass[acmtog, anonymous,review, nonacm, balance = false]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\input{sections/pre.tex}

\newcommand\myworries[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Inverting Colors to save my eye balls
%\usepackage{xcolor}
%\pagecolor[rgb]{0.1,0.12,0.15}
%\color[rgb]{.9,.9,.95}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{acmauthoryear}

\begin{document}

\title{The Shape Matching Element Method: Direct Animation of Curved Surface Models}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{castle.png}
  \caption{tmp}
  \Description{}
  \label{fig:teaser}
\end{teaserfigure}

\begin{abstract}
We introduce a new method for direct physics-based animation of volumetric curved models.  
The Shape Matching Element Method is a completely meshless approach which is robust to gaps and overlaps in geometry, 
is compatible with standard constitutive models and time  integration schemes, supports contact and frictional interactions 
as well as direct editing of simulation output. 
We demonstrate the efficacy of our algorithm by producing compelling physics-based animations from a variety of curved input models.
\end{abstract}

\maketitle

\input{sections/intro}
\input{sections/related}
\input{sections/methods}

\input{sections/results}
\input{sections/conclusion}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}