### 2021-04-27 paper/main.tex
fixing typos

\documentclass[acmtog, anonymous=false, review=false, balance=false]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\input{sections/pre.tex}

\setcopyright{acmlicensed}\acmJournal{TOG}
\acmYear{2021}\acmVolume{40}\acmNumber{4}\acmArticle{69}\acmMonth{8} \acmDOI{10.1145/3450626.3459772}

\acmSubmissionID{230}
\newcommand\myworries[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Inverting Colors to save my eye balls
%\usepackage{xcolor}
%\pagecolor[rgb]{0.1,0.12,0.15}
%\color[rgb]{.9,.9,.95}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{acmauthoryear}

\begin{document}

\title{The Shape Matching Element Method: Direct Animation of Curved Surface Models}

\author{Ty Trusty}

\email{trusty@cs.toronto.edu}
%
\author{Honglin Chen}

\email{chl9797@cs.toronto.edu}
%
%
\author{David I.W. Levin}
\email{diwlevin@cs.toronto.edu}

\affiliation{%
 \institution{University of Toronto}
 \streetaddress{40 St. George Street}
 \city{Toronto}
 \state{ON}
 \country{Canada}
 \postcode{M5S 2E4}
}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10010147</concept_id>
       <concept_desc>Computing methodologies</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10010147.10010371.10010352.10010379</concept_id>
       <concept_desc>Computing methodologies~Physical simulation</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies}
\ccsdesc[500]{Computing methodologies~Physical simulation}

%
% End generated code
%

\keywords{Physically-based animation, NURBS simulation, Meshless simulation, Lagrangian Mechanics}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{figures/teaser_castle.pdf}
  \caption{Using the shape matching element method we can directly simulate this NURBS surface model of a bouncy castle as a volumetric elastic object without the need for volumetric meshing of any kind.}
  \label{fig:teaser}
\end{teaserfigure}

\begin{abstract}
We introduce a new method for direct physics-based animation of volumetric curved models, represented using NURBS surfaces.  
Our technical contribution is the Shape Matching Element Method (SEM).
SEM is a completely meshless algorithm, the first to simultaneously 
be robust to gaps and overlaps in geometry, 
be compatible with standard constitutive models and time integration schemes, support contact and frictional interactions  
and to preserve feature correspondence during simulation which enables editable simulated output. 
We demonstrate the efficacy of our algorithm by producing compelling physics-based animations from a variety of curved input models.
\end{abstract}

\maketitle

\input{sections/intro}
\input{sections/related}
\input{sections/methods}

\input{sections/results}
\input{sections/conclusion}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}